Progression Features & OAC

### 2017-11-03 features-OAC.Rmd
updating glm pipeline

@@

cnv = do.call(rbind, lapply(samples, read.caveman))
totalCNV = merge(cnv, paired, by.x='Illumina.Barcode.full', by.y='SampleId')
oacCNV = makeGRangesFromDataFrame(totalCNV, keep.extra.columns = T)

# Barretts
dir = '~/Data/ICGC/BE_OAC_Cohort/BA/data/ascat'
files = list.files(dir, full.names=T, recursive=T)
samples = list.files(dir, full.names=F)

@@

cnv = do.call(rbind, lapply(samples, read.caveman))
beCNV = merge(cnv, barretts.pairs, by.x='Illumina.Barcode.full', by.y='SampleId')
beCNV = makeGRangesFromDataFrame(beCNV, keep.extra.columns = T)
# ------------------------------ #

@@

table(be.demo$RP.BarrettsAdjacent)
table(oac.demo$RP.BarrettsAdjacent)

@@

sd.pts = do.call(rbind, lapply(swgsPredictions[rownames(sudden)],annotate.by.pt))
bb.pts = do.call(rbind, lapply(swgsPredictions[rownames(born.bad)],annotate.by.pt))
rd.pts = do.call(rbind, lapply(swgsPredictions[sample(which(status == 'P'), 15)],annotate.by.pt))

sd = cbind.data.frame('gain'=apply(rawSWGSdf[subset(sd.pts, !Final)$Samplename,gains$label], 2, count.gains), 
          'loss'=apply(rawSWGSdf[subset(sd.pts, !Final)$Samplename,gains$label], 2, count.losses))

bb = cbind.data.frame('gain'=apply(rawSWGSdf[subset(bb.pts, !Final)$Samplename,gains$label], 2, count.gains), 
          'loss'=apply(rawSWGSdf[subset(bb.pts, !Final)$Samplename,gains$label], 2, count.losses))

rd = cbind.data.frame('gain'=apply(rawSWGSdf[subset(rd.pts, !Final)$Samplename,gains$label], 2, count.gains), 
          'loss'=apply(rawSWGSdf[subset(rd.pts, !Final)$Samplename,gains$label], 2, count.losses))

```
I think this is driven by the number of samples more a difference in the two groups. The "born bad" group has a lot more samples than the "sudden" and if I randomly sample the same number of samples from the "born bad" group the difference disappears. So, if there's a difference here how to I look for it?

```{r}
# Adjusted for the number of samples...is this right?
wilcox.test(sd$gain/nrow(subset(sd.pts, !Final)), bb$gain/nrow(subset(bb.pts, !Final)))
wilcox.test(sd$gain/nrow(subset(sd.pts, !Final)), rd$gain/nrow(subset(rd.pts, !Final)))
wilcox.test(rd$gain/nrow(subset(rd.pts, !Final)), bb$gain/nrow(subset(bb.pts, !Final)))

wilcox.test(sd$loss/nrow(subset(sd.pts, !Final)), bb$loss/nrow(subset(bb.pts, !Final)))
wilcox.test(sd$loss/nrow(subset(sd.pts, !Final)), rd$loss/nrow(subset(rd.pts, !Final)))
wilcox.test(rd$loss/nrow(subset(rd.pts, !Final)), bb$loss/nrow(subset(bb.pts, !Final)))

@@

@@

# one-way anova by ranks -- basically the same as wilcox I think
one.way.aov<-function(a, b) {
  df = rbind(cbind.data.frame('counts'=a, 'class'='a'),
             cbind.data.frame('counts'=b, 'class'='b')) 
  kruskal.test(counts~class, data=df)
  #chisq.test(df$counts, df$class)
  #glm(counts~class, data=df, family='poisson')
}

@@

```{r, warning=F, message=F, fig.width=7}

tests = rbind(cbind.data.frame('CN'='gain', 'p'=wilcox.test(bbG, rrG)$p.value, 'type'='Random', 'label'='BB~R','y'=max(rrG)),
              cbind.data.frame('CN'='gain', 'p'=wilcox.test(bbG, snG)$p.value, 'type'='Sudden', 'label'='BB~SN', 'y'=max(snG)),
              cbind.data.frame('CN'='loss', 'p'=wilcox.test(bbL, rrL)$p.value, 'type'='Random', 'label'='BB~R','y'=max(rrL)),
              cbind.data.frame('CN'='loss', 'p'=wilcox.test(bbL, snL)$p.value, 'type'='Sudden', 'label'='BB~SN', 'y'=max(snL)) )

tmp = rbind(cbind(snG, 'type'='Sudden','CN'='gain'),
            cbind(bbG, 'type'='Born Bad', 'CN'='gain'),
            cbind(rrG, 'type'='Random', 'CN'='gain'),
            cbind(snL, 'type'='Sudden', 'CN'='loss'),
            cbind(bbL, 'type'='Born Bad', 'CN'='loss'),
            cbind(rrL, 'type'='Random', 'CN'='loss' ) )
tmp = as.data.frame(tmp, row.names = NA)
tmp$sn = as.integer( as.character(tmp$sn) )
tmp$type = factor(tmp$type, levels = c('Born Bad','Sudden','Random'))

grid.arrange(
  ggplot(tmp, aes(type, sn, group=type, fill=type)) + facet_grid(~CN) + 
    geom_boxplot(outlier.colour = NA, alpha=0.3) + geom_jitter(width=0.2, aes(color=type)) + 
    #geom_label(data=tests, aes(x=type,y=y,label=paste(label,format.pval(p,2),sep='\np=')), nudge_y=5, alpha=0.5)
    geom_label(data=tests, aes(x=type,y=y,label=paste(label,format.pval(p,2),sep='\np=')), nudge_y=5, alpha=0.5) +
    labs(y='CN count per region',x='', title='Gains & Losses in selected features') + theme(legend.position='none'),

    tableGrob(  rbind(cbind('Group'='Born Bad','# Samples'=length(bb.samples)),
                      cbind('Group'='Sudden','# Samples'=length(sd.samples)), 
                      cbind('Group'='Random','# Samples'=length(random.samples))) ), nrow=2, heights=c(4,1) )

@@

```{r}

@@

m$CN = cut(m$Total_CN, include.lowest=T, ordered_result=T, breaks=c(3, 5, 7, 9, 10, max(m$Total_CN)))
ggplot(data=m, aes(x=order, y=start)) +  
  facet_wrap(~segment, scales='free') +
  geom_segment(aes(x=order, xend=order, y=start, yend=end, color=CN)) + 
  geom_segment(aes(x=0,xend=0, y=seg.start, yend=seg.end), color='red', size=2) +
  scale_colour_brewer(palette='YlGnBu', direction=-1) +
  coord_flip() + 
  theme(legend.position='bottom', axis.text=element_blank(),panel.background=element_blank()) + labs(y='chr position', x='Patient', title='Gains')
}

@@

## gains in patients -- issues with WGD samples?
gains = lapply(swgRegions[which(swgRegions$coef > 0)], gain.for, cnv=oacCNV)
names(gains) = as.character(swgRegions[which(swgRegions$coef > 0)])
plot.gains(gains) + labs(title='Gains in OAC samples')

be.gains = lapply(swgRegions[which(swgRegions$coef > 0)], gain.for, cnv=beCNV)
names(be.gains) = as.character(swgRegions[which(swgRegions$coef > 0)])
plot.gains(be.gains[sapply(be.gains, nrow) > 0]) + labs(title='Gains in BE Adjacent samples')

@@

```{r, fig.height=8, fig.width=8}

@@

gfreq = sort(sapply(be.gains, function(x) length(unique(x$OCCAMS)))/length(unique(be.seq$OCCAMS_ID)))
ggplot(melt(as.matrix(gfreq)), aes(Var1, value)) + geom_col(fill='darkblue')  + coord_flip() + 
  labs(x='', y='Frequency', title='Gain frequency in BE adjacent') + theme_bw() + theme(text=element_text(size=14))

@@

```{r, fig.height=12, fig.width=12}
plot.losses<-function(df) {
  df = lapply(df, function(x) {
    if(nrow(x) <= 0) return(NULL)
    x = arrange(x, -width)
    x$order = factor(1:nrow(x))
    return(x)
  })

  m = do.call(rbind.data.frame, df)
  #m$segnames = ordered(m$seqnames, levels=c(1:22))

  m = m %>% group_by(seqnames) %>% ungroup() %>% arrange(segment, seqnames, width, start) %>% mutate(segorder = row_number())
  m$isWGD = as.factor(m$isWGD)
  m$CN = as.factor(m$Total_CN)
  #m = transform(m, segment=reorder(segment, segnames))
  ggplot(data=m, aes(x=order, y=start)) +  facet_wrap(~segment, scales='free') +
    geom_segment(aes(x=order, xend=order, y=start, yend=end, color=CN)) + 
    geom_segment(aes(x=0,xend=0, y=seg.start, yend=seg.end), color='red', size=2) +
    scale_colour_brewer(palette='YlGnBu', direction=1) +
    coord_flip() + 
    theme(legend.position='bottom', axis.text=element_blank(),panel.background=element_blank()) + labs(y='chr position', x='Patient', title='Losses in OAC Patients')
}

@@

losses = lapply(swgRegions[which(swgRegions$coef < 0)], loss.for, oacCNV)
names(losses) = as.character(swgRegions[which(swgRegions$coef < 0)])
plot.losses(losses)

be.losses = lapply(swgRegions[which(swgRegions$coef < 0)], loss.for, beCNV)
names(be.losses) = as.character(swgRegions[which(swgRegions$coef < 0)])
plot.losses(be.losses)

@@

```{r, fig.height=8, fig.width=8}

lfreq = sort(sapply(losses, function(x) length(unique(x$OCCAMS)))/length(unique(sequenced$OCCAMS_ID)))
ggplot(melt(as.matrix(lfreq)), aes(Var1, value)) + geom_col(fill='darkgreen') + ylim(0,0.5) + coord_flip() + 
  labs(x='', y='Frequency', title='Loss frequency in OAC') + theme_bw() + theme(text=element_text(size=14))

lfreq = sort(sapply(be.losses, function(x) length(unique(x$OCCAMS)))/length(unique(be.seq$OCCAMS_ID)))
ggplot(melt(as.matrix(lfreq)), aes(Var1, value)) + geom_col(fill='darkgreen') + ylim(0,0.5) + coord_flip() + 
  labs(x='', y='Frequency', title='Loss frequency in BE adjacent') + theme_bw() + theme(text=element_text(size=14))

@@

m = arrange(losses$`17:1-4776188`, width, start)
m$Anon = 1:length(m$OCCAMS)
m = transform(m, Anon = reorder(Anon, -width))
segment = cbind('seg.start'=1, 'seg.end'=4776188)
m$CN = as.factor(m$Total_CN)
#ggplot(m, aes(x=Anon, y=start)) +

ggplot() + 
  geom_segment(data=m, aes(x=Anon, xend=Anon, y=start, yend=end, color=CN), size=2) + 
  geom_segment(data=segment, aes(x=0, xend=0, y=seg.start, yend=seg.end),color='red', size=3) + 
  geom_hline(data=chr17, aes(yintercept=(chr.cent-cent.gap)), color='red', linetype='dashed') +
  geom_label(data=chr17, aes(x=as.factor(3), y=chr.cent/2), label='p arm') +
  scale_colour_brewer(palette='YlGnBu', direction=1) +
  coord_flip() + 
  theme(legend.position = 'bottom') + labs(y='chr17 position', x='Patient', title='Chr17 Losses in OAC Patients', subtitle=paste(round(length(unique(losses$`17:1-4776188`$OCCAMS))/length(unique(oacCNV$OCCAMS)), 2)*100, '% of OAC patients', sep=''), text=element_text(size=14, face='bold'))

@@

So is there a difference between the gains/losses in the selected regions?  Not really.
```{r}
wilcox.test(sapply(gains, nrow), sapply(losses, nrow))
one.way.aov(sapply(gains, nrow), sapply(losses, nrow))

# Maybe between BA and tumour...
wilcox.test(sapply(gains, nrow), sapply(be.gains, nrow))
one.way.aov(sapply(gains, nrow), sapply(be.gains, nrow))
```

@@

# Maybe a difference in # gains in the selected features vs random
wilcox.test(sapply(randomGains, nrow), sapply(gains, nrow))
one.way.aov(sapply(randomGains, nrow), sapply(gains, nrow))

@@

# Not sure this shows a really clear difference, but...
df = rbind(cbind.data.frame('counts'=sapply(gains, function(df) length(unique(df$OCCAMS))), 'type'='selected') , 
      cbind.data.frame('counts'=sapply(randomGains, function(df) length(unique(df$OCCAMS))), 'type'='random'))
ggplot(df, aes(type, counts, group=type)) + geom_boxplot() + geom_jitter(width=0.3) + labs(title='Gain counts by region')

df = rbind(cbind.data.frame('counts'=sapply(gains, nrow), 'type'='selected') , 
      cbind.data.frame('counts'=sapply(randomGains, nrow), 'type'='random'))
ggplot(df, aes(type, counts, group=type)) + geom_boxplot() + geom_jitter(width=0.3) + labs(title='Gain counts by patient')

@@

# Definite difference between the numbers of gains and losses in randomly selected regions
wilcox.test(sapply(randomGains, nrow), sapply(randomLosses, nrow))
one.way.aov(sapply(randomGains, nrow), sapply(randomLosses, nrow))

# No difference in # losses in the selected features
wilcox.test(sapply(randomLosses, nrow), sapply(losses, nrow))
one.way.aov(sapply(randomLosses, nrow), sapply(losses, nrow))

df = rbind(cbind.data.frame('counts'=sapply(randomLosses, nrow), 'type'='random') ,
           cbind.data.frame('counts'=sapply(losses, nrow), 'type'='selected'))

#df = df[-which(df$counts > 400),]
# No difference between the two
ggplot(df, aes(type, counts, group=type)) + geom_boxplot() + geom_jitter(width=0.3) + labs(title='Loss counts')

summary(subset(df, type=='random')$counts)
summary(subset(df, type=='selected')$counts)

# No real difference in # patients with losses in the selected features
wilcox.test(sapply(losses, function(df) length(unique(df$OCCAMS))), 
            sapply(randomLosses, function(df) length(unique(df$OCCAMS))))
one.way.aov(sapply(losses, function(df) length(unique(df$OCCAMS))), 
            sapply(randomLosses, function(df) length(unique(df$OCCAMS))))

# No difference between the two
df = rbind(cbind.data.frame('counts'=sapply(losses, function(df) length(unique(df$OCCAMS))), 'type'='selected') , 
      cbind.data.frame('counts'=sapply(randomLosses, function(df) length(unique(df$OCCAMS))), 'type'='random'))
ggplot(df, aes(type, counts, group=type)) + geom_boxplot() + geom_jitter(width=0.3) + labs(title='Loss counts by patient')

#pnorm(subset(df, type=='selected')$counts, mean=mean(subset(df, type=='selected')$counts), sd=sd(subset(df, type=='selected')$counts))
#pnorm(subset(df, type=='random')$counts, mean=mean(subset(df, type=='random')$counts), sd=sd(subset(df, type=='random')$counts))

@@

#  enrichment for some focal changes
p.adjust(apply(sizeGains, 1, function(x) binom.test(x[1], sum(x), 0.5)$p.value ))
apply(sizeGains, 1, function(x) binom.test(x[1], sum(x), 0.5)$conf.int )

# Same is true for gains of type by patient
p.adjust(apply(sizeGainsByPatient, 1, function(x) binom.test(x[1], sum(x), 0.5)$p.value ))
apply(sizeGainsByPatient, 1, function(x) binom.test(x[1], sum(x), 0.5)$conf.int )

>>>>

plot.theme = theme(text=element_text(size=14, face='bold'), panel.background=element_blank(), axis.line=element_line(color='black'), panel.grid.major=element_line(color='grey90'))

@@

cnv = do.call(rbind, lapply(samples, read.caveman))
totalCNV = merge(cnv, paired, by.x='Illumina.Barcode.full', by.y='SampleId')

demo.cols = c('StudySubjectID','RP.TStage.PrimaryTumour','RP.BarrettsAdjacent','Weeks.Survival','RP.TumourDifferentiation','DI.ageAtDiagnosis','DI.PatientGender','FE.Patient.Died')
totalCNV = merge(totalCNV, demo[,demo.cols], by.y='StudySubjectID', by.x='OCCAMS')

oacCNV = makeGRangesFromDataFrame(totalCNV, keep.extra.columns = T)

# Barretts ADJACENT SAMPLES
dir = '~/Data/ICGC/BE_OAC_Cohort/BA/data/ascat'
files = list.files(dir, full.names=T, recursive=T)
samples = list.files(dir, full.names=F)

@@

cnv = do.call(rbind, lapply(samples, read.caveman))
beCNV = merge(cnv, barretts.pairs, by.x='Illumina.Barcode.full', by.y='SampleId')
beCNV = merge(beCNV, demo[,demo.cols], by.y='StudySubjectID', by.x='OCCAMS')
beCNV = makeGRangesFromDataFrame(beCNV, keep.extra.columns = T)
# ------------------------------ #

@@

@@

sd.pts = do.call(rbind, lapply(swgsPredictions[rownames(sudden)],annotate.by.pt))
bb.pts = do.call(rbind, lapply(swgsPredictions[rownames(born.bad)],annotate.by.pt))

# Can't randomly sample patients here because it's very likely that I'll end up with many of the same patients in both random and sudden/born bad groups
# Could randomly select samples that are not HGD/IMC though...
#rd.pts = do.call(rbind, lapply(swgsPredictions[names(status[which(status == 'P')])],annotate.by.pt))

rd.samp = subset(do.call(rbind, swgsPredictions[names(status[which(status == 'P')])]), Pathology %in% c('BE','ID','LGD'), select=c('Prediction','Pathology','Samplename'))
rd.samp = rd.samp[sample(nrow(rd.samp), 120),]

gains.by.group = cbind.data.frame(
  'sudden'=apply(rawSWGSdf[subset(sd.pts, !Final)$Samplename,gains$label], 2, count.gains),
  'born.bad'=apply(rawSWGSdf[subset(bb.pts, !Final)$Samplename,gains$label], 2, count.gains),
  'random'=apply(rawSWGSdf[rd.samp$Samplename,gains$label], 2, count.gains)
)

losses.by.group = cbind.data.frame(
  'sudden'=apply(rawSWGSdf[subset(sd.pts, !Final)$Samplename,losses$label], 2, count.losses),
  'born.bad'=apply(rawSWGSdf[subset(bb.pts, !Final)$Samplename,losses$label], 2, count.losses),
  'random'=apply(rawSWGSdf[rd.samp$Samplename,losses$label], 2, count.losses)
)
```

This may be driven by the number of samples. I'm not sure if this is correct or how to check at the moment.
In all of these I exclude the "final" sample (typically HGD/IMC).

```{r}
# "born bad" and sudden are different
wilcox.test(gains.by.group$sudden, gains.by.group$born.bad)
wilcox.test(losses.by.group$sudden, losses.by.group$born.bad)

# Random and sudden are different
wilcox.test(gains.by.group$sudden, gains.by.group$random)
wilcox.test(losses.by.group$sudden, losses.by.group$random)

# random and born bad aren't?
wilcox.test(gains.by.group$born.bad, gains.by.group$random) ## not so different
wilcox.test(losses.by.group$born.bad, losses.by.group$random) ## not so different

@@

```{r echo=F}
# one-way anova by ranks -- basically the same as wilcox I think
one.way.aov<-function(a, b) {
  df = rbind(cbind.data.frame('counts'=a, 'class'='a'),
             cbind.data.frame('counts'=b, 'class'='b')) 
  kruskal.test(counts~class, data=df)
  #chisq.test(df$counts, df$class)
  #glm(counts~class, data=df, family='poisson')
}
```

So why would random and born bad groups be the same?  Sudden appears to be the only group that stands out in any way, being different from both born bad and random groups.  Sudden has half the samples the other two do, but then random and sudden are different so samplesize should be at issue here.

@@

@@

```{r, warning=F, message=F, echo=F, fig.width=7, fig.height=10}

totals = melt(cbind.data.frame(
  'sudden'=nrow(subset(sd.pts, !Final)),
  'born.bad'=nrow(subset(bb.pts, !Final)),
  'random'=nrow(rd.samp)))

grid.arrange(
  arrangeGrob(
    ggplot(melt(gains.by.group), aes(variable, value, color=variable, fill=variable)) + geom_jitter(width=0.2) + 
    geom_boxplot(alpha=0.5, outlier.color = NA) + labs(title='Gains', x='', y='counts') + theme(legend.position = 'none'),
  ggplot(melt(losses.by.group), aes(variable, value, color=variable, fill=variable)) + geom_jitter(width=0.2) + 
    geom_boxplot(alpha=0.5, outlier.color = NA) + labs(title='Losses', x='', y='counts') + theme(legend.position = 'none'), ncol=2),

  tableGrob(  rbind(cbind('Group'='Born Bad','# Samples'=nrow(subset(bb.pts, !Final))),
                    cbind('Group'='Sudden','# Samples'=nrow(subset(sd.pts, !Final))),
                    cbind('Group'='Random','# Samples'=nrow(rd.samp)))   ) 
  )

@@

@@

m$CN = cut(m$Total_CN, include.lowest=T, ordered_result=T, breaks=c(3, 5, 7, 9, 10, max(m$Total_CN)))
ggplot(data=m, aes(x=order, y=start)) +  
  facet_wrap(~segment, scales='free') +
  geom_segment(aes(x=order, xend=order, y=start, yend=end, color=CN)) + 
  geom_segment(aes(x=0,xend=0, y=seg.start, yend=seg.end), color='red', size=2) +
  scale_colour_brewer(palette='YlGnBu', direction=-1) +
  coord_flip() + plot.theme +
  theme(legend.position='bottom', axis.text=element_blank(),panel.grid.major = element_blank()) + labs(y='chr position', x='Patient', title='Gains')
}

@@

gr = which(swgRegions$coef > 0)

## gains in patients -- issues with WGD samples?
gains = lapply(swgRegions[gr], gain.for, cnv=oacCNV)
names(gains) = as.character(swgRegions[gr])
plot.gains(gains) + labs(title='Gains in OAC samples')

be.gains = lapply(swgRegions[gr], gain.for, cnv=oacCNV[which(oacCNV$RP.BarrettsAdjacent == 'yes')])
names(be.gains) = as.character(swgRegions[gr])
plot.gains(be.gains[sapply(be.gains, nrow) > 0]) + labs(title='Gains in BE Adjacent samples')

non.be.gains = lapply(swgRegions[gr], gain.for, cnv=oacCNV[which(oacCNV$RP.BarrettsAdjacent == 'no')])
names(non.be.gains) = as.character(swgRegions[gr])
plot.gains(non.be.gains[sapply(non.be.gains, nrow) > 0]) + labs(title='Gains in NO BE Adjacent samples')

@@

```{r, fig.height=8, fig.width=12}

@@

gfreq1 = sort(sapply(be.gains, function(x) length(unique(x$OCCAMS)))/length(unique(oacCNV[which(oacCNV$RP.BarrettsAdjacent == 'yes')]$OCCAMS)))
gfreq1 = gfreq1[names(gfreq)]

gfreq2 = sort(sapply(non.be.gains, function(x) length(unique(x$OCCAMS)))/length(unique(oacCNV[which(oacCNV$RP.BarrettsAdjacent == 'no')]$OCCAMS)))
gfreq2 = gfreq2[names(gfreq)]

m = cbind.data.frame('yes'=gfreq1, 'no'=gfreq2)
m$region = rownames(m)
m = transform(m, region=reorder(region, yes))

ggplot(melt(m), aes(region, value, group=variable, fill=variable)) + facet_grid(~variable) + geom_bar(stat='identity', position=position_dodge(), alpha=0.8) + 
  scale_fill_manual(values=c('dodgerblue3','firebrick3'), name='Barretts Adjacent') +
  coord_flip() + labs(x='', y='Frequency', title="Gain frequency BE vs non") + plot.theme + theme(legend.position = 'bottom')

@@

```{r, fig.height=12, fig.width=12}
plot.losses<-function(df) {
  if (!is.data.frame(df)) {
    df = lapply(df, function(x) {
      if(nrow(x) <= 0) return(NULL)
      x = arrange(x, -width)
      x$order = factor(1:nrow(x))
      return(x)
    })
    m = do.call(rbind.data.frame, df)
  } else {
    m = df
    m = arrange(m, -width)
    m$order = factor(1:nrow(m))
  }
  #m$segnames = ordered(m$seqnames, levels=c(1:22))

  m = m %>% group_by(seqnames) %>% ungroup() %>% arrange(segment, seqnames, width, start) %>% mutate(segorder = row_number())
  m$isWGD = as.factor(m$isWGD)
  m$CN = as.factor(m$Total_CN)
  #m = transform(m, segment=reorder(segment, segnames))
  p = ggplot(data=m, aes(x=order, y=start))

  if (length(unique(m$segment)) > 1)
    p = p + facet_wrap(~segment, scales='free')

  p + geom_segment(aes(x=order, xend=order, y=start, yend=end, color=CN)) + 
    geom_segment(aes(x=0,xend=0, y=seg.start, yend=seg.end), color='red', size=2) +
    scale_colour_brewer(palette='YlGnBu', direction=1) +
    coord_flip() + plot.theme +
    theme(legend.position='bottom', axis.text=element_blank(), panel.grid.major=element_blank()) + labs(y='chr position', x='Patient', title='Losses in OAC Patients')
}

@@

lr = which(swgRegions$coef < 0)

losses = lapply(swgRegions[lr], loss.for, oacCNV)
names(losses) = as.character(swgRegions[lr])
plot.losses(losses)

be.losses = lapply(swgRegions[lr], loss.for, oacCNV[which(oacCNV$RP.BarrettsAdjacent == 'yes')])
names(be.losses) = as.character(swgRegions[lr])
plot.losses(be.losses) + labs(title="Losses in (YES) BE Adjacent samples")

non.be.losses = lapply(swgRegions[lr], loss.for, cnv=oacCNV[which(oacCNV$RP.BarrettsAdjacent == 'no')])
names(non.be.losses) = as.character(swgRegions[lr])
plot.losses(non.be.losses) + labs(title='Losses in (NO) BE Adjacent samples')

# grid.arrange(plot.losses(be.losses$`16:76088214-80843726`) + labs(title='BE loss'),
#              plot.losses(non.be.losses$`16:76088214-80843726`) + labs(title='(NO) BE loss') )

@@

```{r, fig.height=8, fig.width=12}

lfreq = sort(sapply(losses, function(x) length(unique(x$OCCAMS)))/length(unique(sequenced$OCCAMS_ID)))
ggplot(melt(as.matrix(lfreq)), aes(Var1, value)) + geom_col(fill='darkgreen') + ylim(0,1) + coord_flip() + 
  labs(x='', y='Frequency', title='Loss frequency in OAC') + theme_bw() + theme(text=element_text(size=14))

lfreq1 = sapply(be.losses, function(x) length(unique(x$OCCAMS)))/length(unique(oacCNV[which(oacCNV$RP.BarrettsAdjacent == 'yes')]$OCCAMS))
lfreq1 = lfreq1[names(lfreq)]

lfreq2 = sort(sapply(non.be.losses, function(x) length(unique(x$OCCAMS)))/length(unique(oacCNV[which(oacCNV$RP.BarrettsAdjacent == 'no')]$OCCAMS)))
lfreq2 = lfreq2[names(lfreq)]

m = cbind.data.frame('yes'=lfreq1, 'no'=lfreq2)
m$region = rownames(m)
m = transform(m, region=reorder(region, yes))

ggplot(melt(m), aes(region, value, group=variable, fill=variable)) + facet_grid(~variable) + geom_bar(stat='identity', position=position_dodge(), alpha=0.8) + 
  scale_fill_manual(values=c('dodgerblue3','firebrick3'), name='Barretts Adjacent') +
  coord_flip() + labs(x='', y='Frequency', title="Loss frequency BE vs non") + plot.theme + theme(legend.position = 'bottom')

```{r fig.height=6,fig.width=8}

@@

plot.losses(losses$`17:1-4776188`) + geom_hline(data=chr17, aes(yintercept=(chr.cent-cent.gap)), color='red', linetype='dashed') +
  labs(y='chr17 position', x='Patient', title='Chr17 Losses in OAC Patients', subtitle=paste(round(length(unique(losses$`17:1-4776188`$OCCAMS))/length(unique(oacCNV$OCCAMS)), 2)*100, '% of OAC patients', sep=''))

@@

So is there a difference between the gains/losses in the selected regions?  
```{r}
wilcox.test(sapply(gains, nrow), sapply(losses, nrow))
one.way.aov(sapply(gains, nrow), sapply(losses, nrow))
```

Not between be/non be either
```{r}
wilcox.test(sapply(non.be.gains, nrow), sapply(be.gains, nrow))
wilcox.test(sapply(non.be.losses, nrow), sapply(be.losses, nrow))
```

@@

# Maybe a difference in # gains in the selected features vs random
wilcox.test(sapply(randomGains, nrow), sapply(gains, nrow))

wilcox.test(sapply(randomGains, nrow), sapply(be.gains,nrow))
wilcox.test(sapply(randomGains, nrow), sapply(non.be.gains,nrow))

@@

df = rbind(cbind.data.frame('counts'=sapply(gains, function(df) length(unique(df$OCCAMS))), 'type'='selected') , 
      cbind.data.frame('counts'=sapply(randomGains, function(df) length(unique(df$OCCAMS))), 'type'='random'))
ggplot(df, aes(type, counts, group=type, fill=type)) + geom_boxplot() + geom_jitter(width=0.3) + labs(title='Patient counts by region') + plot.theme

df = rbind(cbind.data.frame('counts'=sapply(gains, nrow), 'type'='selected') , 
      cbind.data.frame('counts'=sapply(randomGains, nrow), 'type'='random'))
ggplot(df, aes(type, counts, group=type, fill=type)) + geom_boxplot() + geom_jitter(width=0.3) + labs(title='Gain counts per region') + plot.theme

```{r, echo=T}
df = rbind(cbind.data.frame('counts'=sapply(non.be.gains, nrow), 'type'='non.be'),
           cbind.data.frame('counts'=sapply(be.gains, nrow), 'type'='be'),
           cbind.data.frame('counts'=sapply(gains, nrow), 'type'='all.selected'),
           cbind.data.frame('counts'=sapply(randomGains, nrow), 'type'='random'))
ggplot(df, aes(type, counts, group=type, fill=type, color=type)) + geom_boxplot(alpha=0.5, outlier.colour = NA) + 
  geom_jitter(width=0.3) + labs(title='Gain counts per region') + plot.theme

wilcox.test(subset(df, type == 'random')$counts, subset(df, type == 'be')$counts)
wilcox.test(subset(df, type == 'random')$counts, subset(df, type == 'non.be')$counts)
wilcox.test(subset(df, type == 'random')$counts, subset(df, type == 'all.selected')$counts)

@@

# Definite difference between the numbers of gains and losses in randomly selected regions
wilcox.test(sapply(randomGains, nrow), sapply(randomLosses, nrow))

# No difference in # losses in the selected features
wilcox.test(sapply(randomLosses, nrow), sapply(losses, nrow))

wilcox.test(sapply(randomLosses, nrow), sapply(be.losses,nrow))
wilcox.test(sapply(randomLosses, nrow), sapply(non.be.losses,nrow))

# No difference between the two
df = rbind(cbind.data.frame('counts'=sapply(losses, function(df) length(unique(df$OCCAMS))), 'type'='selected') , 
      cbind.data.frame('counts'=sapply(randomLosses, function(df) length(unique(df$OCCAMS))), 'type'='random'))
ggplot(df, aes(type, counts, group=type, fill=type)) + geom_boxplot() + geom_jitter(width=0.3) + labs(title='Loss counts by patient') + plot.theme

df = rbind(cbind.data.frame('counts'=sapply(non.be.losses, nrow), 'type'='non.be'),
           cbind.data.frame('counts'=sapply(be.losses, nrow), 'type'='be'),
           cbind.data.frame('counts'=sapply(losses, nrow), 'type'='all.selected'),
           cbind.data.frame('counts'=sapply(randomLosses, nrow), 'type'='random'))
ggplot(df, aes(type, counts, group=type, fill=type, color=type)) + geom_boxplot(alpha=0.5) + 
  geom_jitter(width=0.3) + labs(title='Loss counts per region') + plot.theme

## Cox

```{r}
library(survival)
library(GGally)

pt.gains = do.call(rbind, lapply(oac.demo$StudySubjectID, function(id) {
  rbind(sapply(gains, function(g)  nrow(subset(g, OCCAMS == id))))
}))
rownames(pt.gains) = oac.demo$StudySubjectID
colnames(pt.gains) = paste('gain_', gsub(':|-', '\\.', colnames(pt.gains)), sep='')

pt.losses = do.call(rbind, lapply(oac.demo$StudySubjectID, function(id) {
  rbind(sapply(losses, function(g)  nrow(subset(g, OCCAMS == id))))
}))
rownames(pt.losses) = oac.demo$StudySubjectID
colnames(pt.losses) = paste('loss_',gsub(':|-', '\\.', colnames(pt.losses)), sep='')

x = merge(pt.gains, pt.losses, by='row.names', all=T)

x = merge(x,oac.demo[,demo.cols], by.x='Row.names', by.y='StudySubjectID')
x$Row.names = NULL

time = 'Weeks.Survival';  event='FE.Patient.Died'; type = '"right"'
x[c(time,event)] = lapply(x[c(time,event)], as.numeric)

x$RP.TStage.PrimaryTumour = as.character(x$RP.TStage.PrimaryTumour)

x[grepl('T1', x$RP.TStage.PrimaryTumour), 'RP.TStage.PrimaryTumour'] = 'T1'
x[grepl('T4', x$RP.TStage.PrimaryTumour), 'RP.TStage.PrimaryTumour'] = 'T4'
x[grepl('Tx', x$RP.TStage.PrimaryTumour), 'RP.TStage.PrimaryTumour'] = 'T0'

terms = as.formula(paste("Surv(",time,",",event,", type=",type,") ~ RP.BarrettsAdjacent", sep=""))
sfit = survfit(terms, data=x,conf.int=0.95, conf.type='log', na.action=na.omit)
ggsurv(sfit)

f1 = paste(demo.cols[c(3,5:7)], collapse='+')
f2 = paste(paste('`', grep('^gain|loss', colnames(x), value=T), '`', sep=''), collapse='+')

f = paste(f1,f2, sep='+')

terms = as.formula(paste("Surv(",time,",",event,", type=",type,") ~", f, sep=""))
cox = coxph(terms, data=x,na.action=na.omit,ties='breslow')
cfit = survfit(cox)
ggsurv(cfit) + labs(title=paste(length(grep('gain|loss', colnames(x))), 'copy number + demo'))

pander(cox)

anova(cox)
```

```{r}
library(CoxHD)

x = x[,c(grep('gain|loss',colnames(x), invert=T), grep('gain|loss',colnames(x)))]
xs = x

surv = with(xs, Surv(Weeks.Survival, FE.Patient.Died))
ggsurv(survfit(surv~1,data=xs))

xs = xs[,-grep('Weeks|Died',colnames(xs))]

xs$RP.BarrettsAdjacent = as.integer(xs$RP.BarrettsAdjacent)-1
xs$RP.TumourDifferentiation = as.integer(as.factor(xs$RP.TumourDifferentiation))-1
xs$DI.PatientGender = as.integer(as.factor(xs$DI.PatientGender))-1

groups = c(rep('clin', 4), rep('cn',(ncol(pt.gains)+ncol(pt.losses)) ))

rfx = CoxRFX(xs, surv, groups, nu=1, verbose=T)
summary(rfx)

plot(rfx)

VarianceComponents(rfx)

WaldTest(rfx)

PartialC(rfx)

PlotVarianceComponents(rfx)

@@

#  enrichment for some focal changes
x = which(p.adjust(apply(sizeGains, 1, function(x) binom.test(x[1], sum(x), 0.5)$p.value )) < 0.05)
apply(sizeGains, 1, function(x) binom.test(x[1], sum(x), 0.5)$conf.int )[,x]

## Not sure what to do with this really
sizeGainsBE = do.call(rbind.data.frame, lapply(be.gains, countBySize))
which(p.adjust(apply(sizeGainsBE, 1, function(x) binom.test(x[1], sum(x), 0.5)$p.value )) < 0.05)
sizeGainsNonBE = do.call(rbind.data.frame, lapply(non.be.gains, countBySize))
which(p.adjust(apply(sizeGainsNonBE, 1, function(x) binom.test(x[1], sum(x), 0.5)$p.value )) < 0.05)
